A Method for the Spectroscopic Inference of Stellar Parameters

### 2014-08-19 tex/paper.tex
set up for kurucz
lots of minor Sect 1, 2 edits; first iteration of Sect 3 edits

%kernel notation
\newcommand{\KK}{\mathcal{K}}
\newcommand{\Kglobal}{\KK_{\textrm{g}}}
\newcommand{\Klocal}{\KK_l}

\newcommand{\todo}[1]{ \textcolor{blue}{\\TODO: #1}}
\newcommand{\comm}[1]{ \textcolor{red}{SA: #1}}

@@

\title{A Method for the Spectroscopic Inference of Stellar Parameters}
\author{Ian Czekala, Sean M.~Andrews, et al.}
\affil{Harvard-Smithsonian Center for Astrophysics, 60 Garden Street, Cambridge, MA 02138}
%\email{iczekala@cfa.harvard.edu}

@@

There are various well-tested techniques being used to compare these models with observed spectra 
and thereby infer basic stellar parameters; here we highlight three examples to illustrate some key 
ideas.  First is a straightforward empirical approach that relies on distilling an information-rich 
subset of the spectral data, usually in the form of spectral line equivalent widths and/or local 
continuum shapes.  A combined sequence of the ratios of these quantities are found to be especially 
sensitive to one key parameter in the models, and therefore can be used to quantify it 
\citep[e.g.,][]{gray94,reid95,rojas-ayala10,rojas-ayala12}.  The line index comparison approach has 
the major advantage of being trivially fast, but each index relationship is only informative over 
a limited range of parameter-space.  A second technique exploits the cross-correlation of an 
observed spectrum with a suite of model templates to identify an optimized set of parameters, 
usually with some preferential weighting applied to specific spectral regions \citep[e.g., {\tt 
SPC};][]{buchhave12}.  In this case, the speed advantage is maintained while taking advantage of 
more data content (particularly in the spectral dimension), thereby achieving high precision even 
for data with comparatively low sensitivity; the disadvantage is that the model quality and 
parameter inferences are assessed in an empirical, rather than probabilistic, framework.  The third 
approach of interest prefers a direct, pixel-by-pixel comparison between model and observed 
spectra, and is usually associated with a spectral synthesis front-end (e.g., {\tt MOOG}; 
\citealt{sneden73}; {\tt SME}; \citealt{valenti96}).  This technique has the benefits of parametric 
flexibility (e.g., one can fit for atomic abundances or arbitrary structures) and a proper 
inference framework \citep[usually a least-squares approach, although increasingly in a Bayesian 
format;][]{shkedy07,schoenrich13}, but it is computationally expensive and can be biased by model 
systematics \citep[e.g.,][]{mann13}; in practice, this often requires an effort focused on a 
pre-selected subsample of the data.

In this article, we design a flexible forward-modeling approach to the general spectroscopic 
inference problem in a Bayesian framework, building on the best aspects of the latter two methods 
highlighted above.  Employing a non-trivial covariance matrix parameterized by both global 
(stationary) and local (non-stationary) Gaussian process kernels, we are able to account for 
residual pixel-by-pixel correlations that arise when comparing observed spectra to intrinsically 
imperfect models.  This approach minimizes the potential for bias, efficiently propagates 
systematic uncertainties into the parameter inferences, and ultimately can serve as tool for using
observations to guide substantially improvements in the models themselves.  An overview of the 
methodology of this approach is provided in Section \ref{sec:method}, including the covariance 
matrix formalism.  Some relevant tests and example applications (specifically for a high resolution 
optical spectrum of an F star, and a medium-resolution near-infrared spectrum of a mid-M star) of 
the method are described in Section \ref{sec:examples}.  Finally, a discussion of the potential 
utility of the technique, and the possibility of extending it to develop data-driven spectral 
models, is provided in Section \ref{sec:discussion}.

@@

Here we describe a generative Bayesian modeling framework that confronts some of the key obstacles 
in the spectroscopic inference problem.  The goal of this approach is to conservatively extract the 
maximal amount of information about a prescribed (and degenerate) parameter set by forward-modeling 
an observed spectrum, while also recognizing and explicitly accounting for the covariances and 
biases introduced by pathologically imperfect models or calibrations.  The method is modular, and 
therefore can easily incorporate additional physical or nuisance parameters as desired without 
sacrificing an accurate reflection of the limitations in the data.  Moreover, with a well-crafted 
observational sample, this data-driven approach should ultimately enable us to systematically learn 
how synthetic spectral models can be improved.  The specific applications discussed here are 
related to the spectra of individual stars, but the methodology is generic (and could be used for 
composite spectra of unresolved stellar clusters, galaxies, etc.).

The remainder of this section describes the mechanics of this modeling framework.  First, a model 
spectrum is generated for a given set of physical parameters (Section \ref{subsec:synthetic}), and 
then post-processed to mimic reality using a set of observational and practical nuisance parameters 
(Section \ref{subsec:postprocess}).  Next, a direct, pixel-by-pixel comparison between the data and 
model spectra is made with a prescribed likelihood function and a parametric treatment of the 
covariances between pixel residuals (Section \ref{subsec:likelihood}).  That process is iterated 
under the guise of hierarchical Markov Chain Monte Carlo (MCMC) simulations to numerically explore 
the posterior probability density of the model conditioned on the data, and thereby to determine 
constraints on the parameters of interest (Section \ref{subsec:MCMC}).  Along the way, these 
procedures are illustrated with real observations of the high resolution optical spectrum from a 
nearby F star.  That specific application, along with some alternative demonstrations of the 
method, are discussed in more detail in Section \ref{sec:examples}.

@@

There are various approaches to synthesizing a spectrum, $f_{\lambda}$, for a specific set of model
parameters, $\vt_{\ast}$.  In an ideal case, a model stellar atmosphere is constructed and then 
subsequently processed through a radiative transfer code \citep[e.g.,][]{kurucz93,hauschildt99}.  
However, in general this approach is still computationally prohibitive for any iterative method of 
probabilistic inference.  One partial compromise is to interpolate over a library of atmosphere 
structures that were pre-computed for a discrete grid of parameter values, $\{\vt_{\ast}\}^{\rm 
grid}$, for some arbitrary $\vt_{\ast}$, and then perform a radiative transfer calculation with 
that interpolated atmosphere to synthesize $\flam$ \citep[e.g., as for {\tt SME};][]{valenti96}.  A 
more common variant is to instead rely on interpolation over a library of pre-synthesized model 
spectra, $\flam(\vt_{\ast})$ \citep[e.g.,][]{castelli04,allard12,husser13}.  While technically the 
former approach is most similar to the ideal case, the computational cost of repeated spectral 
synthesis is sufficiently high to make a detailed exploration of parameter space (particularly for 
data with a large spectral range) considerably less appealing.  A related, but different approach 
is to eschew forward modeling entirely (and therefore repeated spectral syntheses and/or library 
interpolations), and instead evaluate the models only at the discrete grid points of the library.  
Then, these discretized samples of the posterior probability density can be interpolated to an 
arbitrary $\vt_{\ast}$ to construct appropriate confidence intervals \citep[similar to the method 
of {\tt SPC};][]{buchhave12}.  The difficulty with this latter approach is that the parameter 
uncertainties can be smaller than the grid spacing; in that case, there is valid concern that this 
interpolation might not accurately recover intrinsic parameter degeneracies.

Here we opt to take the computationally expedient approach that employs a library of model spectra, 
$\flam(\{\vt_{\ast}\}^{\rm grid})$, where $\vt_{\ast} = [T_{\rm eff}, \,\, \log{g}, \,\, Z]$ 
(in practice, the metallicity $Z$ is often parameterized by [Fe/H]).  However, it is worth noting 
that the techniques we will develop are applicable to {\it any} reasonable ``front-end" that 
generates a model spectrum.  In our adopted approach, the model spectrum for an arbitrary 
$\vt_{\ast}$ must be interpolated from among the spectral library, 
\begin{equation}
\flam(\{\vt_{\ast}\}^{\rm grid}) \leadsto \flam(\vt_{\ast}),
\label{eqn:interp}
\end{equation}
where we assign the symbol $\leadsto$ as an interpolation operator.  The multi-dimensional 
interpolation in Eq.~\ref{eqn:interp} needs to be performed many times, so computational efficiency 
is critical.  In practice, a simple tri-linear interpolation is suitably fast, but introduces an 
undesireable level of inaccuracy (particularly in the $Z$ dimension).  The interpolation quality 
can be empirically estimated by performing the operation in Eq.~\ref{eqn:interp} across a 
calculated location in $\{\vt_{\ast}\}^{\rm grid}$, and then comparing the interpolated spectrum 
with the corresponding library spectrum.  After an extensive exploration of such calculations 
\citep[see also][]{husser12}, we concluded that the best combination of speed and accuracy can be 
achieved by pre-computing a {\it refined} spectral library using a cubic spline interpolation with 
a $\{\vt_{\ast}\}^{\rm grid}$ spacing of [20\,K, 0.1\,dex, 0.1\,dex], and then performing 
tri-linear interpolation over that refined grid.  Overall, this interpolation technique is found to 
be accurate within a few percent per high resolution model pixel.  Ideally, this pre-interpolation 
could be avoided if the spectral library was computed over a refined grid (with a substantial 
up-front computational investment); but for the time being, we can empirically propagate these 
interpolation uncertainties into the likelihood calculations (as will be described in 
Section~\ref{subsec:likelihood}).  
\todo{Talk about tests to show that it is within \% accuracy}.

@@

Generally, the ``raw" model spectrum $\flam(\vt_{\ast})$ will be highly over-sampled compared to a 
typical observed spectrum, and does not account for several additional observational and 
instrumental effects that become important in comparisons with real data.  Therefore, a certain 
amount of post-processing is required before assessing the model quality.  We treat that 
post-processing in two stages: the first deals with an additional set of ``observational" 
parameters, $\vt_{\rm obs}$, that incorporate dynamical effects, geometry, and the relative 
location of the target, while the second employs a suite of nuisance (hyper-)parameters ($\vp$) 
designed to mitigate an imperfect data calibration.

We can further divide $\vt_{\rm obs}$ into those parameters that impact the model primarily in the 
spectral or flux dimensions; $\vt_{\rm obs} = [\vt_{\rm obs, v}, \,\, \vt_{\rm obs, f}]$.  For the 
former, we consider three kernels that contribute to the observed line-of-sight velocity 
distribution function, $\varphi_v$.  The first, $\mathcal{F}_v^{\rm inst}$,
treats the instrumental spectral broadening\footnote{For computational
efficiency, we first pre-convolve the entire raw spectral library with a slight
under-prediction of $\mathcal{F}_v^{\rm inst}$ and resample the
spectra to this lower resolution. Because the instrumental PSF acts as a
low-pass filter, this procedure can reduce the size of the library by a factor
of 10 or more while also preserving the final Fourier content of the
spectrum. Then, when a given spectrum is post-processed, the actual $\mathcal{F}_v^{\rm
inst}$ used is a correction to the original
under-prediction.}.  For illustrative purposes we assume
$\mathcal{F}_v^{\rm inst}$ is a Gaussian 
with a mean at $v = 0$ and a constant width $\sigma_v$ at all $\lambda$, although more 
sophisticated forms could be adopted.  The second, $\mathcal{F}_v^{\rm rot}$, characterizes the 
broadening induced by (projected) stellar rotation, parameterized by $v\sin{i}$ as described by 
\citet[][his Eq.~18.14]{gray08}.  And the third, $\mathcal{F}_v^{\rm dop} = \delta(v-v_z)$, 
incorporates the radial velocity through a Doppler shift.  The model spectrum is modified by the 
parameters $\vt_{{\rm obs}, v} = [\sigma_v, \,\, v\sin{i}, \,\, v_z]$ through these kernels, using 
a convolution in velocity-space,\footnote{In practice, these convolutions are performed as 
multiplications in Fourier-space to better preserve spectral information \citep[cf.,][]{tonry79}; 
the mathematical formalism is presented here for clarity.}
\begin{equation}
\flam(\vt_{\ast}, \vt_{{\rm obs}, v}) = \flam(\vt_{\ast}) \otimes \varphi_v = \flam(\vt_{\ast}) \otimes \mathcal{F}_v^{\rm inst} \otimes \mathcal{F}_v^{\rm rot} \otimes \mathcal{F}_v^{\rm dop},
\label{eqn:broadening}
\end{equation}
and then re-sampled onto the discrete wavelengths corresponding to each data pixel, 
\begin{equation}
\flam(\vt_{\ast}, \vt_{{\rm obs}, v}) \mapsto \vM(\vt_{\ast}, \vt_{{\rm obs}, v}),
\label{eqn:resampling}
\end{equation}
where the $\mapsto$ symbol denotes a re-sampling operator that maps the model spectrum onto the 
$N_{\rm pix}$-element model vector $\vM$ (and $N_{\rm pix}$ is the number of pixels in the 
spectrum).  For reference, Figure \ref{fig:broadening} shows a graphical representation of these 
post-processing steps.

\begin{figure}[!t]
\begin{center}
  \includegraphics{figs/kernels.pdf}
  \includegraphics{figs/high2low.pdf}
  \caption{Panel 1: the instrumental (${\rm FWHM} =6.8\;\kms$) and rotational ($v \sin i = 5\;\kms$) broadening kernels, and their composite. Panel 2: A section of the raw, high resolution synthetic model at $R=$500,000. Panel 3: The spectrum after applying the composite broadening kernel, $\varphi_v$. Panel 4: The broadened spectrum downsampled to the pixels of the detector.}
\label{fig:broadening}
\end{center}
\end{figure}

At this stage, the model is further modified in the flux dimension.  A typical synthetic spectrum 
is computed as the flux that would be measured {\it at the stellar surface}, and so needs to be 
diluted by the subtended solid angle, $\Omega = (R_{\ast}/d)^2$, where $R_{\ast}$ is the stellar 
radius and $d$ is the distance.  An additional wavelength-dependent scaling factor is applied to 
account for interstellar extinction, assuming some previously-derived extinction law $A_{\lambda}$ 
\citep[e.g.,][]{cardelli89} that is parameterized by $A_V$.  The parameters $\vt_{{\rm obs}, f} = 
[\Omega, \,\, A_V]$ are applied as
\begin{equation}
\vM(\vT) \equiv \vM(\vt_{\ast}, \vt_{{\rm obs},v}, \vt_{{\rm obs},f}) = \vM(\vt_{\ast}, \vt_{{\rm obs}, v}) \times \Omega \times 10^{-0.4\,A_{\lambda}},
\label{eqn:scaling}
\end{equation}
where we have simplified the notation by writing $\vT \equiv [\vt_{\ast}, \,\, \vt_{\rm obs}]$.

So far, the procedure summarized in Eq.~\ref{eqn:interp}-\ref{eqn:scaling} is composed of 
relatively straightforward operations demanded by practical astronomical and computing issues.  If 
the observed spectrum was {\it perfectly} calibrated, we could proceed to a likelihood calculation 
(Section~\ref{subsec:likelihood}) that makes a direct comparison with $\vM(\vT)$ at this point.  
However, that is unlikely to be the case.  The primary concern is that an imperfect calibration 
produces (presumably low-level) mismatches in the underlying shape of the observed spectrum on 
relatively broad wavelength scales.  When compared with the model, these mismatches might represent 
a non-trivial contribution to the likelihood, and thereby bias our estimates of the desired 
physical parameters.  Often, this concern is treated externally to any modeling procedure, usually 
by dividing the observed spectrum (and model spectrum) by an appropriate polynomial function.  But 
that ``normalization" procedure implicitly assumes that there is no relevant information content on 
those mismatched scales; if $\vT$ also contributes to the broad spectral shape, then adopting this 
approach will corrupt the inferences of these parameters.  Moreover, in practice this approach is 
limited, since defining an appropriate polynomial becomes difficult in cases where the spectral 
line density is high (e.g., molecular bands for cool stars).

We adopt a somewhat analogous approach to deal with this issue, but cast it {\it internal} to the 
modeling framework to appropriately propagate the uncertainty introduced by additional degrees of 
freedom in the model.  The residual calibration errors are treated as an explicit contribution to 
the model spectrum, enabling us to explore the distribution of possible calibrations (see 
Section~\ref{subsec:MCMC}) and then eventually marginalize out the associated nuisance parameters.  
In essence, the inferences for the relevant stellar parameters will properly account for the 
underlying uncertainty in the calibration process.  In practice, this is achieved by distorting 
segments of the model spectrum with polynomials, $\mathsf{P}$ \citep[e.g.,][]{eisenstein06,
koleva09}.  For data with $N_{\rm ord}$ spectral orders, each denoted with index $o$ (where $o$ 
naturally has an implicit $\lambda$-dependence), we can write
\begin{equation}
\vM(\vT, \cheb) = \vM(\vT) \times \{\mathsf{P}_o\} \equiv \vM(\vT) \times \left\{ \sum_n c_o^{(n)} \, T_o^{(n)} \right\},
\label{eqn:chebyshev}
\end{equation}
where $T^{(n)}$ are the standard Chebyshev functions of degree $n$ and the coefficients are treated 
as a set of $n \times N_{\rm ord}$ nuisance (hyper-)parameters, $\cheb = [\{\vp_{\mathsf{P}_o}\}] = 
[\{c_o^{(0)}, c_o^{(1)}, \ldots, c_o^{(n-1)} \}]$.  With judicious priors on $\cheb$, we can ensure 
that the unintended treatment of real spectral features (e.g., broad, deep molecular bands) as 
calibration artifacts is negligible (see Section~\ref{sec:examples} for examples).  The 
lowest-degree (scaling) coefficient, $c^{(0)}$, is by its nature degenerate with the solid angle 
parameter, $\Omega$.  Therefore, we enforce an additional constraint by requiring that the mean of 
the polynomial is unity.  For data with a single spectral order, this means simply setting $c^{(0)} 
= 1$.  If the goal is to model multiple spectral orders, we assign $c^{(0)} = 1$ in an arbitrary 
order as an anchor, but permit the $c^{(0)}$ values in other orders to be different (as 
necessary).  It is worth noting that this formalism can, in principle, be extended to develop 
models for completely uncalibrated spectra, rather than the residual calibration mismatches as 
described here (with a suitable relaxation of the priors on $\cheb$).  Figure \ref{fig:chebyshev} 
offers a practical demonstration of how these nuisance parameters are applied.

>>>>

%kernel notation
\newcommand{\KK}{\mathcal{K}}
\newcommand{\Kglobal}{\KK^{\textrm{g}}}
\newcommand{\Klocal}{\KK^l}

\newcommand{\todo}[1]{ \textcolor{blue}{\\TODO: #1}}
\newcommand{\comm}[1]{ \textcolor{red}{SA: #1}}
\newcommand{\hili}[1]{ \textcolor{green}{#1}}

@@

\title{A Method for the Spectroscopic Inference of Stellar Parameters}
\author{Ian Czekala, Sean M.~Andrews, et al.}
\affil{Harvard-Smithsonian Center for Astrophysics, 60 Garden Street, Cambridge, MA 02138}
\email{iczekala@cfa.harvard.edu}

\begin{abstract}
An abstract needs to go here. \\
tbd \\
tbd \\
tbd \\
tbd \\
tbd \\
tbd \\
tbd \\
tbd \\
tbd \\
tbd \\
tbd
\end{abstract}
\keywords{stars: fundamental parameters --- methods: data analysis --- methods: statistical}

@@

There are various well-tested techniques being used to compare these models with observed spectra 
and thereby infer basic stellar parameters; here we highlight three examples to illustrate some key 
ideas.  First is a straightforward empirical approach that relies on distilling an information-rich 
subset of the data, usually in the form of spectral line equivalent widths and/or local continuum 
shapes.  A combined sequence of the ratios of these quantities are found to be especially sensitive 
to one key parameter in the models, and therefore can be used to quantify it 
\citep[e.g.,][]{gray94,reid95,rojas-ayala10,rojas-ayala12}.  This index comparison approach has 
the advantage of being trivially fast, but each index relationship is only informative over a 
limited region of parameter-space.  A second technique exploits the cross-correlation of an 
observed spectrum with a suite of model templates to identify an optimized set of parameters, 
usually with some weighting applied to specific spectral regions \citep[e.g., {\tt 
SPC};][]{buchhave12}.  In this case, the speed advantage is maintained and more data content is 
used (particularly in the spectral dimension), thereby achieving higher precision even for data 
with comparatively low sensitivity; the disadvantage is that the model quality and parameter 
inferences are assessed in an empirical, rather than probabilistic, framework.  A third approach 
employs a direct, pixel-by-pixel comparison between model and data, and is usually associated with 
a spectral synthesis front-end (e.g., {\tt MOOG}; \citealt{sneden73}; {\tt SME}; 
\citealt{valenti96}).  This technique has the benefits of parametric flexibility (e.g., one can fit 
for arbitrary abundances or structures) and a proper inference framework \citep[usually a 
least-squares approach, although increasingly in a Bayesian format;][]{shkedy07,schoenrich13}, but 
it is computationally expensive and can be biased by systematics \citep[e.g.,][]{mann13}; in 
practice, this often requires an effort focused on a pre-selected subsample of the data.

In this article, we design a flexible forward-modeling approach to the general spectroscopic 
inference problem in a Bayesian framework, building on the best aspects of the latter two methods 
highlighted above.  Employing a non-trivial covariance matrix parameterized by both global 
(stationary) and local (non-stationary) Gaussian process kernels, we are able to account for 
residual pixel-to-pixel correlations that arise when comparing observed spectra to intrinsically 
imperfect models.  This approach minimizes the potential for bias, efficiently propagates 
systematic uncertainties into the parameter inferences, and ultimately can serve as tool for using
observations to guide substantially improvements in the models themselves.  An overview of the 
methodology of this approach is provided in Section \ref{sec:method}, including the covariance 
matrix formalism.  Some tests and example applications (for a high resolution optical spectrum of 
an F star, and a medium-resolution near-infrared spectrum of a mid-M star) of the method are 
described in Section \ref{sec:examples}.  Finally, a discussion of the potential utility of the 
technique, and the possibility of extending it to develop data-driven spectral models, is provided 
in Section \ref{sec:discussion}.  \\

@@

In this Section, we describe a generative Bayesian modeling framework that confronts some of the 
key obstacles in the spectroscopic inference problem.  The goal of this approach is to 
conservatively extract the maximal amount of information about a prescribed (and usually 
degenerate) parameter set by forward-modeling an observed spectrum, while also recognizing and 
explicitly accounting for the covariances and biases introduced by pathologically imperfect 
models.  The method is modular, and therefore can easily incorporate additional physical or 
nuisance parameters as desired without sacrificing an accurate reflection of the limitations in the 
data.  Moreover, with a well-crafted observational sample, this data-driven approach should 
ultimately enable us to systematically learn how synthetic spectral models can be improved.  The 
specific applications discussed here are related to the spectra of individual stars, but the 
methodology is generic (and could be used for the composite spectra of unresolved stellar clusters, 
galaxies, etc.).

The remainder of this Section describes the mechanics of this modeling framework.  First, a model 
spectrum is generated for a given set of physical parameters (Section \ref{subsec:synthetic}), and 
then post-processed to mimic reality using a set of observational and practical nuisance parameters 
(Section \ref{subsec:postprocess}).  Next, a direct, pixel-by-pixel comparison between the data and 
model spectra is made with a prescribed likelihood function and a parametric treatment of the 
covariances between pixel residuals (Section \ref{subsec:likelihood}).  That process is iterated 
under the guise of hierarchical Markov Chain Monte Carlo (MCMC) simulations to numerically explore 
the posterior probability density of the model conditioned on the data, and thereby to determine 
constraints on the parameters of interest (Section \ref{subsec:MCMC}).  Along the way, these 
procedures are illustrated with real observations of the high resolution optical spectrum from a 
nearby F star.  That specific application, along with some alternative demonstrations of the 
method, are discussed in more detail in Section \ref{sec:examples}.

@@

There are various approaches to synthesizing a spectrum, $f_{\lambda}$, for a specific set of model
parameters, $\vt_{\ast}$.  In an ideal case, a model stellar atmosphere is constructed and then 
subsequently processed through a radiative transfer code \citep[e.g.,][]{kurucz93,hauschildt99}.  
However, in general this approach is still computationally prohibitive for any iterative method of 
probabilistic inference.  One partial compromise is to interpolate over a library of atmosphere 
structures that were pre-computed for a discrete grid of parameter values, $\{\vt_{\ast}\}^{\rm 
grid}$, for some arbitrary $\vt_{\ast}$, and then perform a radiative transfer calculation with 
that interpolated atmosphere to synthesize $\flam$ \citep[e.g., as for {\tt SME};][]{valenti96}.  A 
more common variant is to instead rely on interpolation over a library of pre-synthesized model 
spectra, $\flam(\vt_{\ast})$ \citep[e.g.,][]{castelli04,allard12,husser13}.  While technically the 
former approach is most similar to the ideal case, the computational cost of repeated spectral 
synthesis is sufficiently high to make a detailed exploration of parameter space (particularly for 
data with a large spectral range) considerably less appealing.  An alternative approach eschews 
forward modeling entirely (and therefore repeated spectral syntheses and/or library 
interpolations), and instead evaluates the models only at the discrete grid points of the library.  
Then, these discretized samples of the posterior probability density can be interpolated to an 
arbitrary $\vt_{\ast}$ to construct appropriate confidence intervals \citep[similar to the method 
of {\tt SPC};][]{buchhave12}.  The difficulty with this latter approach is that the parameter 
uncertainties can be smaller than the grid spacing; in that case, there is valid concern that this 
interpolation might not accurately recover intrinsic parameter degeneracies.

We opt for the computationally expedient approach that employs a library of model spectra, 
$\flam(\{\vt_{\ast}\}^{\rm grid})$, where $\vt_{\ast} = [T_{\rm eff}, \,\, \log{g}, \,\, 
[{\rm Fe/H}]]$.  However, it is worth noting that the techniques developed here are applicable to 
{\it any} ``front-end" that generates a model spectrum.  In our adopted approach, the model 
spectrum for an arbitrary $\vt_{\ast}$ is interpolated from a spectral library, 
\begin{equation} \label{eqn:interp}
\flam(\{\vt_{\ast}\}^{\rm grid}) \leadsto \flam(\vt_{\ast}),
\end{equation}
where we assign the symbol $\leadsto$ as an interpolation operator.  The multi-dimensional 
interpolation in Eq.~\ref{eqn:interp} needs to be performed many times, so computational efficiency 
is critical.  In practice, a simple tri-linear interpolation is suitably fast, but introduces an 
undesirable level of inaccuracy (particularly in the [Fe/H] dimension).  The interpolation quality 
can be empirically estimated by performing the operation in Eq.~\ref{eqn:interp} across a 
$\{\vt_{\ast}\}^{\rm grid}$ grid point, and then comparing the interpolated spectrum with the 
corresponding library spectrum.  After an extensive exploration of such calculations 
\citep[see also][]{husser12}, we concluded that the best combination of speed and accuracy can be 
achieved by pre-computing a {\it refined} spectral library using a cubic spline interpolation with 
a $\{\vt_{\ast}\}^{\rm grid}$ spacing of [20\,K, 0.1\,dex, 0.1\,dex], and then performing 
tri-linear interpolation over that refined grid.  Overall, this interpolation technique is found to 
be accurate within a few percent per full-resolution model pixel.  Ideally, this pre-interpolation 
could be avoided if the spectral library was computed over a refined grid (with a substantial 
up-front computational investment); but for the time being, we can empirically propagate these 
interpolation uncertainties into the likelihood calculations (as will be described in 
Section~\ref{subsec:likelihood}).  
%% \todo{tests regarding interpolation accuracy}.

@@

Generally, the ``raw" model spectrum $\flam(\vt_{\ast})$ is highly over-sampled compared to the 
observed spectrum, and does not account for several additional observational and 
instrumental effects that become important in comparisons with real data.  Therefore, a certain 
amount of post-processing is required before assessing the model quality.  We treat that 
post-processing in two stages: the first deals with an additional set of ``observational" 
parameters, $\vt_{\rm obs}$, that incorporate dynamical effects, geometry, and the relative 
location of the target, while the second employs a suite of nuisance (hyper-)parameters ($\vp$) 
designed to mitigate an imperfect data calibration.

We can further divide $\vt_{\rm obs}$ into those parameters that impact the model primarily in the 
spectral or flux dimensions.  For the former, we consider three kernels that contribute to the 
line-of-sight velocity distribution function, $\varphi_v$.  The first, $\mathcal{F}_v^{\rm inst}$, 
treats the instrumental spectral broadening.\footnote{For computational efficiency, we first 
pre-convolve the entire raw spectral library with a slight under-prediction of $\mathcal{F}_v^{\rm 
inst}$ and resample the spectra to a lower resolution.  Because the instrumental point spread 
function acts as a low-pass filter, this procedure reduces the size of the library by a factor of 
10 or more while still preserving the final Fourier content of the spectra.  Then, when any given 
spectrum is post-processed, the actual $\mathcal{F}_v^{\rm inst}$ used is treated as a smaller 
``correction" to this pre-processing kernel.}  For illustrative purposes, we assume 
$\mathcal{F}_v^{\rm inst}$ is a Gaussian with mean $v = 0$ and constant width $\sigma_v$ at all 
$\lambda$, although more sophisticated forms could be adopted.  The second, $\mathcal{F}_v^{\rm 
rot}$, characterizes the broadening induced by (projected) stellar rotation, parameterized by 
$v\sin{i}$ as described by \citet[][his Eq.~18.14]{gray08}.  And the third, $\mathcal{F}_v^{\rm 
dop} = \delta(v-v_z)$, incorporates the radial velocity through a Doppler shift.  The model 
spectrum is modified by the parameters $[\sigma_v, \,\, v\sin{i}, \,\, v_z]$ through these kernels, 
using a convolution in velocity-space,\footnote{In practice, these convolutions are performed as 
multiplications in Fourier-space to better preserve spectral information \citep[cf.,][]{tonry79}; 
the mathematical formalism is presented for clarity.}
\begin{eqnarray} \label{eqn:broadening}
\flam(\vt_{\ast}, \sigma_v, v\sin{i}, v_z) &=& \flam(\vt_{\ast}) \otimes \varphi_v \\
                                           &=& \flam(\vt_{\ast}) \otimes \mathcal{F}_v^{\rm inst} \otimes \mathcal{F}_v^{\rm rot} \otimes \mathcal{F}_v^{\rm dop}, \nonumber
\end{eqnarray} 
and then re-sampled onto the discrete wavelengths corresponding to each data pixel, 
\begin{equation} \label{eqn:resampling}
\flam(\vt_{\ast},  \sigma_v, v\sin{i}, v_z) \mapsto \vM(\vt_{\ast},  \sigma_v, v\sin{i}, v_z),
\end{equation}
where the $\mapsto$ symbol denotes a re-sampling operator that maps the model spectrum onto the 
$N_{\rm pix}$-element model vector $\vM$ (and $N_{\rm pix}$ is the number of pixels in the 
spectrum).  Figure \ref{fig:broadening} shows a (condensed) graphical representation 
of these post-processing steps.

\begin{figure}[!t]
\begin{center}
  \includegraphics{figs/kernels.pdf}
  \includegraphics{figs/high2low.pdf}
  \figcaption{({\it top}) The line-of-sight velocity distribution function, $\varphi_v$, and its 
decomposition into broadening kernels.  The instrumental kernel ({\it dotted}) is treated as 
a Gaussian, the rotation kernel ({\it dashed}) is a Gaussian-like function of the projected 
rotational velocity, and the Doppler kernel ({\it solid}) is a $\delta$-function that introduces 
the radial velocity.  In this specific case, $\sigma_v = 2.9$\,km s$^{-1}$, $v\sin{i} = 5$\,km 
s$^{-1}$, and $v_z = 7$\,km s$^{-1}$, appropriate for the example in Section \ref{sec:examples}.x.
({\it bottom}) A segment of a raw, full-resolution model spectrum and its post-processed equivalent 
after convolution by $\varphi_v$ and re-sampling at the coarser resolution of the detector pixels. 
\label{fig:broadening}}
\end{center}
\end{figure}

At this stage, the model is further modified in the flux dimension.  A typical synthetic spectrum 
is computed as the flux that would be measured {\it at the stellar surface}, and so needs to be 
diluted by the subtended solid angle, $\Omega = (R_{\ast}/d)^2$, where $R_{\ast}$ is the stellar 
radius and $d$ is the distance.  An additional wavelength-dependent scaling factor is applied to 
account for interstellar extinction, assuming some previously-derived extinction law $A_{\lambda}$ 
\citep[e.g.,][]{cardelli89} that is parameterized by $A_V$.  The parameters $[\Omega, \,\, A_V]$ 
are then applied as
\begin{eqnarray} \label{eqn:scaling}
\vM(\vT) &=& \vM(\vt_{\ast}, \vt_{\rm obs}) \\
         &=& \vM(\vt_{\ast}, \sigma_v, v\sin{i}, v_z) \times \Omega \times 10^{-0.4\,A_{\lambda}}, \nonumber
\end{eqnarray}
with simplified notation such that $\vT \equiv [\vt_{\ast}, \,\, \vt_{\rm obs}]$, 
where $\vt_{\rm obs} = [\sigma_v, v\sin{i}, v_z, \Omega, A_V]$.

The procedure so far, summarized in Eq.~\ref{eqn:interp}--\ref{eqn:scaling}, is composed of 
straightforward operations demanded by practical astronomical and computing issues.  If the data 
were {\it perfectly} calibrated, we could proceed to a likelihood calculation that makes a direct 
comparison with $\vM(\vT)$.  However, that is unlikely to be the case.  The key concern is that an 
imperfect calibration might produce mismatches in the underlying shapes of the data and model.  
Although such mismatches presumably occur at a low-level over relatively broad wavelength scales, 
they still might represent a non-trivial likelihood contribution, and thereby bias our estimates of 
the physical parameters.  This concern is treated externally to any modeling procedure, most often 
by dividing the observed spectrum (and model spectrum) by a polynomial in $\lambda$.  But that 
``normalization" implicitly assumes that there is no relevant information content in the spectral 
shape; if $\vT$ also contributes on the mismatched scales, then adopting this approach will corrupt 
the inferences of these parameters.  Moreover, in practice this approach is limited, since defining 
an appropriate polynomial becomes difficult in cases where the spectral line density is high (e.g., 
molecular bands for cool stars).

We employ an analogous, but more rigorous, approach to deal with this issue, cast {\it internal} to 
the modeling framework to propagate the uncertainties introduced by additional degrees of freedom 
in the model.  Deviations in the spectral shape introduced by calibration residuals are treated as 
explicit contributions to the model spectrum, enabling an exploration of the distribution of 
possible ``tweaks" to the calibration that can be compartmentalized into a set of nuisance 
parameters and then marginalized.  This way, the inferences for the stellar parameters will 
properly account for the underlying uncertainty in the calibration of the spectral shape.  In 
practice, this is achieved by distorting segments of the model spectrum with polynomials, 
$\mathsf{P}$ \citep[e.g.,][]{eisenstein06,koleva09}.  For data that has $N_{\rm ord}$ spectral 
orders, each denoted with index $o$, the model spectrum can be decomposed into the set of orders
\begin{eqnarray} \label{eqn:chebyshev}
\vM(\vT, \cheb) &=& \{ \vM_o(\vT) \times \mathsf{P}_o \} \\
                &=& \{ \vM_o(\vT) \times \sum_n c_o^{(n)} \, T_o^{(n)} \}, \nonumber
\end{eqnarray}
and $T^{(n)}$ is the $n^{\rm th}$ degree Chebyshev function.  The $n \times N_{\rm ord}$ 
coefficients are considered a set of nuisance (hyper-)parameters, $\cheb = [\{c_o^{(0)}, c_o^{(1)}, 
\ldots, c_o^{(n-1)} \}]$.  With judicious priors, we can ensure that the unintended treatment of 
real spectral features (e.g., molecular bands) as residual calibration artifacts is negligible.  
The lowest-degree (scaling) coefficient, $c^{(0)}$, is naturally degenerate with the solid angle, 
$\Omega$.  Therefore, we enforce an additional constraint that the mean of the polynomial is 
unity.  For data with a single spectral order, this means simply setting $c^{(0)} = 1$.  In the 
multiple order case, we assign $c^{(0)} = 1$ in an arbitrary order as an anchor, but permit the 
$c^{(0)}$ in other orders to be different.  It is worth noting that this formalism can, in 
principle, be extended to develop models for completely uncalibrated spectra, rather than the 
residual calibration mismatches described here (with a suitable relaxation of the priors on 
$\cheb$).  Figure \ref{fig:chebyshev} offers a practical demonstration of how these nuisance 
parameters are applied.
